Slitasje og kjørespor

### 2022-12-22 DRAFT_slitasjeOgKjørespor.Rmd
style: text edits

<!-- Replace 'indicator name' with yor own indicator name  -->

@@

<br />

@@

<br />

<!-- Load all you dependencies here -->

@@

<!-- Fill in which ecosystem the indicator belongs to, as well as the ecosystem characteristic it should be linked to. It's OK to use some Norwegian here -->

@@

<!-- Don't remove these three html lines -->

<br /> <br />

<hr />

<!-- Document you work below. Try not to change  the headers too much. Data can be stored on NINA server. Since the book is rendered on the R Server this works fine, but note that directory paths are different on the server compared to you local machine. If it is not too big you may store under /data/ on this repository -->

@@

```{r}
naturetypes <- readRDS("/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/rds/naturtyper.rds")

@@

Calculate area

@@

The nin variables are combined into a single field which we need to spit
apart.

@@

Convert the value column to numeric. Ignore NA's

@@

Keep only the relevant variables

@@

Colum names starting with a number is problematic, so adding a prefix

@@

Remove the old PR variables

@@

This dataset only contains data from year 2019 and 2021.

@@

```{r}
par(mar=c(5,6,4,2))
plot(ano$andel_hovedoekosystem_punkt[order(ano$andel_hovedoekosystem_punkt)],
     ylab="Percentage of the 250 m2 area\ncovered by the main ecosystem")
```

@@

Let's now look at the distribution of the variables. First I need to
sepertae the variable name fro the values themsolves. Now the data looks
like this:

@@

I can reduce the number of columns to keep things a bit more tidy

```{r}
temp <- as.data.frame(ano)
temp2 <- as.data.frame(naturetypes)
slitasje_data <- full_join(select(temp,
                                  GlobalID,
                                  origin,
                                  kartleggingsår,
                                  hovedøkosystem,
                                  area,
                                  slitasje,
                                  SHAPE), 
                           select(temp2,
                                  identifikasjon_lokalid,
                                  origin,
                                  hovedøkosystem,
                                  kartleggingsår,
                                  area,
                                  slitasje,
                                  SHAPE))
# convert back to sf object
slitasje_data <- st_as_sf(slitasje_data)
st_crs(slitasje_data) <- myCRS
```

>>>>

@@

@@

@@

@@

<hr/>

@@

```{r}
path <- ifelse(dir == "C:", 
               "P:/41201785_okologisk_tilstand_2022_2023/data/rds/naturtyper.rds",
               "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/rds/naturtyper.rds")
naturetypes <- readRDS(path)
```

@@

Calculate area for each polygon.

@@

The NiN variables are combined into a single field which we need to spit
apart.

@@

Convert the value column to numeric. Ignore NA's.

@@

Keep only the relevant variables.

@@

Names starting with a number is problematic, so adding a prefix

@@

Remove the old PR variables.

@@

This dataset only contains data from year 2019 and 2021. We need to
update this dataset later.

@@

```{r, fig.cap="Distribution of the ANO variable andel_hovedoekosystem_punkt for wetland localities."}
par(mar=c(5,6,4,2))
plot(ano$andel_hovedoekosystem_punkt[order(ano$andel_hovedoekosystem_punkt)],
     ylab="Percentage of the 250 m2 area\ncovered by the main ecosystem")
```

@@

Let's now look at the distribution of the variables. First I need to
seperate the variable name fro the values themselves. Now the data looks
like this:

@@

I can reduce the number of columns to keep things a bit more tidy.

<!--# this join doesnt work! -->

```{r, eval=F}
temp <- as.data.frame(ano)
temp2 <- as.data.frame(naturetypes)
slitasje_data <- full_join(select(temp,
                                  GlobalID,
                                  origin,
                                  kartleggingsår,
                                  hovedøkosystem,
                                  area,
                                  slitasje,
                                  SHAPE), 
                           select(temp2,
                                  identifikasjon_lokalid,
                                  origin,
                                  hovedøkosystem,
                                  kartleggingsår,
                                  area,
                                  slitasje,
                                  SHAPE))
# convert back to sf object
slitasje_data <- st_as_sf(slitasje_data)
st_crs(slitasje_data) <- myCRS
```
